---
output: html_document
---
Title: "Reproducible Research: Peer Assessment 1"
==================================================================

## Loading and preprocessing the data
The data are loaded here, please put the "activity.csv" in your working directory
```{r YangCode1, echo=TRUE, results='asis'}
library(ggplot2)
library(lubridate)
activity <- read.table("activity.csv", head=T, sep=",", stringsAsFactors=F)
```
## What is mean total number of steps taken per day
```{r meanCalculation, echo=TRUE}
total_day <- tapply (activity$steps, activity$date, sum)
mean_totoal_day <- mean (total_day, na.rm=T)
median_total_day <- median(total_day, na.rm=T)
```
The mean total number of steps taken per day is `r mean_totoal_day`, median is `r median_total_day`.
The histogram of total steps per day
```{r}
hist(total_day)
```

## What is the average daily activity pattern?
```{r average, results='hide'}
summary.steps_1<-aggregate(activity$steps, list(activity$interval), na.rm=T, mean)
summary.steps <- data.frame(mean.steps=summary.steps_1[,2], uni.interval=summary.steps_1[,1])
ggplot(summary.steps, aes(y=mean.steps, x=uni.interval))+geom_line()
```
Calculate the interval which contains the maximal average of steps
```{r}
max_steps <- max (summary.steps[,1])
index=which(summary.steps[,1]==max_steps)
max <- summary.steps[index,2]
```
The interval containing maximal steps is 'r max'.

## Imputing missing values
The missing values were filled in with average of the corresponding interval of every day 
```{r fill in missing value}
activity_new <- activity
for (i in 1:length(activity_new$steps)){
    if (is.na(activity_new[i,1])){
        activity_new[i,1] <- summary.steps_1[which(summary.steps_1[,1]==activity_new[i,3]),2]
    }    
}
```
## Are there differences in activity patterns between weekdays and weekends?
Convert time variable
```{r}
activity_new[,4] <- ymd(activity_new[,2])
names(activity_new)[4] <- "data.time"
activity_new[,5] <- character(0)
names(activity_new)[5] <- "weekdays"
```
Add new column "weekday"
```{r}
for (i in 1:length(activity_new$data.time)){
    if (weekdays(activity_new[i,4]) %in% c("Saturday", "Sunday")){
        activity_new[i,5] <- "weekend"
    }
    else {
        activity_new[i,5] <- "weekday" 
    }
}
```
Calculate activity pattern in weekday and weekends separately
```{r}
summary.steps_weekday<-aggregate(activity_new[activity_new[,5]=="weekday",]$steps, list(activity_new[activity_new[,5]=="weekday",]$interval), na.rm=T, mean)
summary.steps_weekday[,3] <- "weekday"
str(summary.steps_weekday)
summary.steps_weekend<-aggregate(activity_new[activity_new[,5]=="weekend",]$steps, list(activity_new[activity_new[,5]=="weekend",]$interval), na.rm=T, mean)
summary.steps_weekend[,3] <- "weekend"
str(summary.steps_weekend)
summary.steps_combined <- rbind (summary.steps_weekday,summary.steps_weekend)

summary.steps_weekday_weekend <- data.frame(mean.steps=summary.steps_combined[,2], interval=summary.steps_combined[,1], day=summary.steps_combined[,3])
```
Plot the result with ggplot2
```{r}
ggplot(summary.steps_weekday_weekend, aes(y=mean.steps, x=interval))+geom_line()+facet_grid(day~.)
```

Based on the plot, the activity patterns are different between workdays and weekends. 




